\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning}
\usepackage[pdfusetitle]{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue,
            urlcolor=blue}

% Alonzo macros (course-provided)
\input{simple-type-theory-def.tex}
\input{alonzo-notation.tex}

% Theory name macros
\newcommand{\PERM}{\mName{PERM}}
\newcommand{\FILETYPE}{\mName{FILE}}
\newcommand{\SECFILE}{\mName{SECURED\text{-}FILE}}
\newcommand{\DIR}{\mName{DIRECTORY}}
\newcommand{\DIRDEV}{\mName{DIR\text{-}DEV}}

\newcommand{\FILEPERM}{\mName{FILE\text{-}PERM}}

% Constructor/operation name macros
\newcommand{\mkfile}{\mName{mkfile}}
\newcommand{\fname}{\mName{fname}}
\newcommand{\fcontent}{\mName{fcontent}}
\newcommand{\perm}{\mName{perm}}
\newcommand{\prot}{\mName{protect}}

\title{CAS 760 --- Development 2 Project:\\
       A Simple File System in Alonzo}
\author{Suleyman Kiani \and Hamza Jamil}
\date{April 16, 2026}

\begin{document}
\maketitle
\tableofcontents
\newpage

% ============================================================
\section{Introduction}
\label{sec:intro}
% ============================================================

This paper formalizes a simple file system as a development graph
in Alonzo, following the little theories method~\cite{farmer2025}.
File systems are a fundamental concept in computing: they organize
data into named, content-bearing objects (files), assign permissions
to control access, and group files into directories.  Each of these
concepts corresponds naturally to a theory in the little theories
framework.

We build a hierarchy of increasingly complex structures.  The base
theories $\PERM$ and $\FILETYPE$ capture permissions and basic files
independently.  The theory extension $\SECFILE$ combines both,
associating a permission level with each file.  A further extension
$\DIR$ introduces directories as inductive structures over secured
files.  Key operations (lookup, containment, directory size) are
developed in $\DIRDEV$, and a theory morphism shows that single files
embed into the directory structure.

The development follows the module format established in the textbook
and the monoids paper of Farmer \& Zvigelsky~\cite{monoids2025}.
We use compact notation throughout, dropping type subscripts where
unambiguous, consistent with Notational Convention~5
of~\cite{farmer2025}.

\medskip
\noindent\textbf{Theory graph.}

\begin{center}
\begin{tikzpicture}[
  theory/.style={draw, rounded corners, minimum width=2.2cm,
                 minimum height=0.7cm, font=\small\ttfamily},
  dev/.style={draw, dashed, rounded corners, minimum width=2.2cm,
              minimum height=0.7cm, font=\small\ttfamily},
  arr/.style={-{Stealth[length=6pt]}, thick},
  morph/.style={-{Stealth[length=6pt]}, thick, dashed},
  node distance=1.4cm
]
  \node[theory] (perm)  {PERM};
  \node[theory, right=3cm of perm] (file) {FILE};
  \node[theory, below right=1.4cm and 0.5cm of perm] (secfile) {SECURED-FILE};
  \node[theory, below=of secfile] (dir) {DIRECTORY};
  \node[dev, below left=1cm and 0cm of dir] (dirdev) {DIR-DEV};

  \draw[arr] (perm) -- (secfile);
  \draw[arr] (file) -- (secfile);
  \draw[arr] (secfile) -- (dir);
  \draw[arr] (dir) -- (dirdev);
  \draw[morph] (file) to[bend right=30]
    node[right, font=\scriptsize] {morphism} (dir);
\end{tikzpicture}
\end{center}

Solid boxes are theories (theory definitions or extensions).  The
dashed box is a development definition.  Solid arrows denote theory
extensions.  The dashed arrow denotes the theory morphism
$\mName{FILE\text{-}to\text{-}DIR}$.

\medskip

% ============================================================
\newpage
\section{Permissions}
\label{sec:perm}
% ============================================================

A file system associates \emph{permissions} with files to control
access.  We model a simple permission hierarchy with four levels:
no access, read-only, read-write, and administrator.  These form
an enumeration inductive type --- a finite set freely generated by
four nullary constructors, analogous to the theory of Boolean
values (BOOLE) in Chapter~10 of~\cite{farmer2025}.

\begin{theory-def}
{Permissions}
{$\PERM$.}
{$P$.}
{$\mName{noaccess}_P$,\;
 $\mName{rd}_P$,\;
 $\mName{wr}_P$,\;
 $\mName{admin}_P$.}
{
\be
\item $\cDistinctL{\mName{noaccess},\, \mName{rd},\,
      \mName{wr},\, \mName{admin}}$
      \hfill (all constructors are distinct).

\item $\cForallX{p}{\cSetTy{P}}{
      \cImpliesX{p\;\mName{noaccess} \;\wedge\;
        p\;\mName{rd} \;\wedge\;
        p\;\mName{wr} \;\wedge\;
        p\;\mName{admin}}{
        \cForallX{x}{P}{p\;x}}}$
      \hfill (structural induction).
\ee
}
\end{theory-def}

\medskip
\noindent\textbf{Explanation.}
Axiom~1 states that the four permission values are pairwise
distinct (using the $\mName{DISTINCT}$ abbreviation from
Table~6.2 of~\cite{farmer2025}).  Axiom~2 is the structural
induction principle: any predicate that holds for all four
constructors holds for every element of $P$.  Together, these
axioms make $(P, K_P)$ an enumeration inductive type where
$K_P = \{\mName{noaccess},\; \mName{rd},\; \mName{wr},\;
\mName{admin}\}$, so in every standard model, $P$ has exactly
four elements.  (We write $K_P$ for the set of constructors of $P$,
following the standard inductive-type convention; the subscript
also keeps the notation distinct from the base type $C$ of
$\FILETYPE$.)

% ============================================================
\newpage
\section{Files}
\label{sec:files}
% ============================================================

A \emph{file} is a named container for content.  In a real file
system, a file has (at minimum) a name and content.  We model
this as a theory with a constructor $\mkfile$ that builds a file
from a name and content, and accessor functions $\fname$ and
$\fcontent$ that extract the components.

This theory defines the space of all possible files over arbitrary
name and content types.  It is analogous to a simple record type
$\texttt{type file = \{name : N; content : C\}}$ in OCaml.

\begin{theory-def}
{Files}
{$\FILETYPE$.}
{$N$, $C$, $F$.}
{$\mkfile_{\cFunTyBX{N}{C}{F}}$,\;
 $\fname_{\cFunTyX{F}{N}}$,\;
 $\fcontent_{\cFunTyX{F}{C}}$.}
{
\be
\item $\cTotalB{\mkfile}$
      \hfill ($\mkfile$ is total).

\item $\cInjB{\mkfile}$
      \hfill ($\mkfile$ is injective).

\item $\cForallBX{n}{N}{c}{C}{
      \cEqX{\fname\;(\mkfile\;n\;c)}{n}}$
      \hfill (name accessor).

\item $\cForallBX{n}{N}{c}{C}{
      \cEqX{\fcontent\;(\mkfile\;n\;c)}{c}}$
      \hfill (content accessor).

\item $\cTotal{\fname}$
      \hfill ($\fname$ is total).

\item $\cTotal{\fcontent}$
      \hfill ($\fcontent$ is total).

\item $\cForallX{f}{F}{\cEqX{\mkfile\;(\fname\;f)\;(\fcontent\;f)}{f}}$
      \hfill (surjectivity of $\mkfile$).
\ee
}
\end{theory-def}

\medskip
\noindent\textbf{Explanation of the axioms.}
Axiom~1 states that any name--content pair produces a valid file.
Axiom~2 ensures distinct pairs produce distinct files (``no
confusion'').  Axioms~3--4 define the accessor functions via their
interaction with the constructor.  Axioms~5--6 ensure the accessors
are total on all files.  Axiom~7 establishes that every file is
constructible --- i.e., $\mkfile$ is surjective, so the type $F$ is
exactly the set of name--content pairs.  Together, Axioms~1--2 and~7
make $F$ isomorphic to $N \times C$.

\medskip
\noindent\textbf{Design note.}
We could alternatively model files using the product type
$N \times C$ with the built-in projection functions $\mName{fst}$
and $\mName{snd}$.  We choose an abstract type $F$ with explicit
accessors to allow extension with additional attributes (e.g.,
permissions) via theory extensions, which is more natural in the
little theories setting.

% ============================================================
\newpage
\section{Secured Files}
\label{sec:secfiles}
% ============================================================

A \emph{secured file} is a file that has been assigned a permission
level.  Conceptually, $\SECFILE$ combines $\FILETYPE$ and $\PERM$:
it associates a permission from $\PERM$ with each file from
$\FILETYPE$.  We present $\SECFILE$ as a joint extension of both
parent theories, so that the base type $P$ is inherited from
$\PERM$ (not reintroduced), and the base types $N$, $C$, $F$ and
the constants $\mkfile$, $\fname$, $\fcontent$ are inherited from
$\FILETYPE$.  This is the standard pushout-style construction for
combining two theories that share no non-trivial sub-theory,
following the little theories method of~\cite{farmer2025}: the
two extension arrows $\FILETYPE \to \SECFILE$ and $\PERM \to
\SECFILE$ in the theory graph of Section~\ref{sec:intro} encode
the import of each parent's language into $\SECFILE$.  $\SECFILE$
then introduces only the new constants $\perm$ and $\prot$,
together with axioms relating them to the inherited constants.

\begin{theory-ext}
{Secured Files}
{$\SECFILE$.}
{$\FILETYPE$ and $\PERM$.}
{None ($P$ is inherited from $\PERM$).}
{$\perm_{\cFunTyX{F}{P}}$,\;
 $\prot_{\cFunTyBX{F}{P}{F}}$.}
{
\be \setcounter{enumi}{7}
\item $\cTotal{\perm}$
      \hfill ($\perm$ is total).

\item $\cTotalB{\prot}$
      \hfill ($\prot$ is total).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\perm\;(\prot\;f\;p)}{p}}$
      \hfill (protect sets permission).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\fname\;(\prot\;f\;p)}{\fname\;f}}$
      \hfill (protect preserves name).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\fcontent\;(\prot\;f\;p)}{\fcontent\;f}}$
      \hfill (protect preserves content).
\ee
}
\end{theory-ext}

\medskip
\noindent\textbf{Explanation of the axioms.}
Axiom~8 states that every file has an assigned permission.
Axiom~9 ensures that applying a permission always succeeds.
Axiom~10 defines the core semantics of $\prot$: after applying
$\prot\;f\;p$, the permission of the resulting file is $p$.
Axioms~11--12 guarantee that $\prot$ only changes the permission
--- it does not alter the file's name or content.

$\SECFILE$ inherits the types and constants of both $\FILETYPE$
and $\PERM$ (including the base type $P$ of permission levels) and
adds two new constants, $\perm$ and $\prot$.  The axiom numbering
continues from $\FILETYPE$ (Axioms 1--7), consistent with the
convention for theory extensions (Section~9.2
of~\cite{farmer2025}).  The language of $\SECFILE$ is the union of
the languages of $\FILETYPE$ and $\PERM$, augmented with the
constants $\perm$ and $\prot$.


%----------------

\newpage
\section{Secured Files}
\label{sec:secfiles}
% ============================================================

A \emph{secured file} is a file that has been assigned a permission
level.  Formally, we construct the theory of secured files in two
steps, following the theory-graph combinators of Chapter~14
of~\cite{farmer2025}.

First, let $\FILEPERM$ be the theory obtained by combining
$\FILETYPE$ and $\PERM$ over a common trivial subtheory, as in the
theory combination construction of Example~14.63
of~\cite{farmer2025}.  Since $\FILETYPE$ and $\PERM$ have disjoint
base types and disjoint constants, this combination is defined and
yields a theory whose language is the union of the languages of
$\FILETYPE$ and $\PERM$.  Thus $\FILEPERM$ contains the base types
$N$, $C$, $F$, $P$ and the constants
$\mkfile$, $\fname$, $\fcontent$,
$\mName{noaccess}$, $\mName{rd}$, $\mName{wr}$, and
$\mName{admin}$, together with all the axioms of both parent
theories.

Second, we define $\SECFILE$ as an ordinary theory extension of
$\FILEPERM$, following the theory extension construction of
Example~14.61 of~\cite{farmer2025}.  The only new symbols added are
$\perm$ and $\prot$, which relate the inherited permission type $P$
to the inherited file type $F$.

\begin{theory-ext}
{Secured Files}
{$\SECFILE$.}
{$\FILEPERM$.}
{None.}
{$\perm_{\cFunTyX{F}{P}}$,\;
 $\prot_{\cFunTyBX{F}{P}{F}}$.}
{
\be
\item $\cTotal{\perm}$
      \hfill ($\perm$ is total).

\item $\cTotalB{\prot}$
      \hfill ($\prot$ is total).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\perm\;(\prot\;f\;p)}{p}}$
      \hfill (protect sets permission).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\fname\;(\prot\;f\;p)}{\fname\;f}}$
      \hfill (protect preserves name).

\item $\cForallBX{f}{F}{p}{P}{
      \cEqX{\fcontent\;(\prot\;f\;p)}{\fcontent\;f}}$
      \hfill (protect preserves content).
\ee
}
\end{theory-ext}

\medskip
\noindent\textbf{Explanation of the axioms.}
Axiom~1 states that every file has an assigned permission.
Axiom~2 ensures that applying a permission always succeeds.
Axiom~3 defines the core semantics of $\prot$: after applying
$\prot\;f\;p$, the permission of the resulting file is $p$.
Axioms~4--5 guarantee that $\prot$ only changes the permission
--- it does not alter the file's name or content.

The theory $\SECFILE$ is therefore obtained by extending the
combined theory $\FILEPERM$ with the constants $\perm$ and $\prot$
and the five axioms above.  Its language contains all the symbols of
$\FILETYPE$ and $\PERM$, together with the new constants
$\perm$ and $\prot$.
%------------------
% ============================================================
\newpage
\section{Directories}
\label{sec:dir}
% ============================================================

A \emph{directory} is a collection of secured files organized as a
flat structure.  We model this as an inductive type: a directory is
either empty or formed by adding a secured file to an existing
directory.  This mirrors the standard inductive structure of lists
(cf.\ the parallel between PA and lists in Theory
Definition~9.7 of~\cite{farmer2025}).

We add two new constructors: $\mName{emptydir}$ (the empty
directory, analogous to $\mName{nil}$ in lists) and $\mName{addfile}$ (add
a file to a directory, analogous to $\mName{cons}$).

\medskip
\noindent\textbf{Design note on flat structure.}
The constructor $\mName{addfile}$ has type
$\cFunTyBX{F}{D}{D}$: it takes a \emph{file} (not a directory) as
its first argument.  This makes $\DIR$ a flat container of
$\SECFILE$ entries, mirroring the LIST pattern in the textbook.
A hierarchical model would require a constructor
$\mName{adddir} : \cFunTyBX{D}{D}{D}$ allowing directories to
contain directories, yielding a tree-shaped inductive type.  The
flat choice is intentional: it keeps the inductive type within the
standard fourth-kind constructor pattern of Chapter~10
of~\cite{farmer2025}, preserves the direct analogy with lists, and
permits the clean $\mName{FILE\text{-}to\text{-}DIR}$ morphism in
Section~\ref{sec:morphism}, which relies on unique extraction from
single-entry directories.  A hierarchical extension is a natural
direction for future work.

\begin{theory-ext}
{Directories}
{$\DIR$.}
{$\SECFILE$.}
{$D$.}
{$\mName{emptydir}_D$,\;
 $\mName{addfile}_{\cFunTyBX{F}{D}{D}}$.}
{
\be \setcounter{enumi}{12}
\item $\cTotalB{\mName{addfile}}$
      \hfill ($\mName{addfile}$ is total).

\item $\cInjB{\mName{addfile}}$
      \hfill ($\mName{addfile}$ is injective).

\item $\cForallBX{f}{F}{d}{D}{
      \cNotEqX{\mName{addfile}\;f\;d}{\mName{emptydir}}}$
      \hfill (disjointness).

\item $\cForallX{p}{\cSetTy{D}}{
      \cImpliesX{p\;\mName{emptydir} \;\wedge\;
        \cForallBX{f}{F}{d}{D}{
          \cImpliesX{p\;d}{p\;(\mName{addfile}\;f\;d)}}}{
        \cForallX{d}{D}{p\;d}}}$
      \hfill (structural induction).
\ee
}
\end{theory-ext}

\medskip
\noindent\textbf{Inductive type verification.}
The constructors for $D$ are
$K_D = \{\mName{emptydir}_D,\;
  (\mName{addfile}_{\cFunTyBX{F}{D}{D}},\; U_{\cSetTy{F}})\}$.
Here $\mName{emptydir}$ is a nullary constructor (first kind) and
$(\mName{addfile}, U_{\cSetTy{F}})$ is a fourth-kind constructor:
its type is $F \to D \to D$ (in Curryed form), the quasitype
$U_{\cSetTy{F}}$ indicates the first argument ranges over all
of~$F$, and the presence of $D$ in the domain makes
$\mName{addfile}$ recursive.

Checking the obligations (Chapter~10, Section~10.2
of~\cite{farmer2025}):
\begin{enumerate}[label=(\alph*)]
  \item $U_{\cSetTy{F}} \ne \emptyset_{\cSetTy{F}}$: holds since
    $F$ is a base type and every base type domain is nonempty
    (frame condition F2, Definition~5.1 of~\cite{farmer2025}).
  \item $\mName{addfile}\!\downarrow$ as a total function:
    $\cTotalB{\mName{addfile}}$ (Axiom~13).
  \item TOTAL-ON for $\mName{addfile}$: follows from Axiom~13.
  \item $\cInjB{\mName{addfile}}$: Axiom~14.  This ensures
    distinct file--directory pairs produce distinct directories.
  \item $\mName{emptydir} \notin \mName{ran}(\mName{addfile})$:
    follows from Axiom~15 (disjointness).
  \item Disjoint ranges: trivially satisfied (only one nonnullary
    constructor).
  \item Structural induction: Axiom~16.
\end{enumerate}

Hence $(D, K_D)$ is a recursive inductive type in $\DIR$.
By the Inductive Types Theorem (Chapter~10 of~\cite{farmer2025}),
in every standard model $M$ of $\DIR$, $D_D^M$ consists of exactly
$I^M(\mName{emptydir})$ and all values of the form
$I^M(\mName{addfile})(d_f, \sigma)$ for $d_f \in D_F^M$ and
$\sigma \in D_D^M$.  That is, $D_D^M$ is isomorphic to the set of
finite lists of elements of $D_F^M$.

\medskip
\noindent\textbf{Comparison with LIST-TYPE.}
The following table shows the structural parallel between
$\DIR$ and a list type:

\medskip
\begin{center}
\begin{tabular}{lll}
  \textbf{LIST} & \textbf{DIRECTORY} & \textbf{Role} \\
  \hline
  $L$ & $D$ & inductive type \\
  $E$ (element) & $F$ (file) & element type \\
  $\mName{nil}_L$ & $\mName{emptydir}_D$ & nullary constructor \\
  $\mName{cons}_{E \to L \to L}$ & $\mName{addfile}_{F \to D \to D}$ &
    recursive constructor \\
  TOTAL2($\mName{cons}$) & TOTAL2($\mName{addfile}$) & totality \\
  $\mName{cons}\;e\;l \ne \mName{nil}$ & $\mName{addfile}\;f\;d \ne \mName{emptydir}$
    & disjointness \\
  INJ2($\mName{cons}$) & INJ2($\mName{addfile}$) & injectivity \\
\end{tabular}
\end{center}

% ============================================================
\newpage
\section{Directory Development}
\label{sec:dir-dev}
% ============================================================

We now define derived operations on directories and prove key
properties.  Following the lesson from Development Project~1
(see Section~\ref{sec:conclusion}), all operations are
\emph{defined} in this development using definite descriptions,
not introduced as theory extension constants.

\begin{dev-def}
{Directory Development}
{$\DIRDEV$.}
{$\DIR$.}
{

\medskip
\noindent\textbf{Def\,1} ($\mName{contains}$). \quad
$\mName{contains}_{\cFunTyBX{N}{D}{\cB}} \;=\;
\cDefDesX{g}{N \to D \to \cB}{
  \cForallX{n}{N}{\cEqX{g\;n\;\mName{emptydir}}{\cF}} \;\wedge\;
  \cForallCX{n}{N}{f}{F}{d}{D}{
    \cEqX{g\;n\;(\mName{addfile}\;f\;d)}{
      (\cEqX{\fname\;f}{n}) \vee g\;n\;d}}}$

\medskip
\noindent\textit{Remark.}
The predicate $\mName{contains}\;n\;d$ tests whether a file with
name $n$ exists in directory $d$.  It is defined by structural
recursion on~$d$: the empty directory contains nothing, and
$\mName{addfile}\;f\;d$ contains name $n$ if either $f$ has
name $n$ or $d$ already contains $n$.

\medskip
\noindent\textbf{Thm\,1.} \quad $\cTotalB{\mName{contains}}$
\hfill ($\mName{contains}$ is total).

\medskip
\noindent\textbf{Thm\,2.} \quad
$\cForallX{n}{N}{\cEqX{\mName{contains}\;n\;\mName{emptydir}}{\cF}}$
\hfill (empty directory contains nothing).

\medskip
\noindent\textbf{Def\,2} ($\mName{size}$). \quad
$\mName{size}_{\cFunTyX{D}{\mathbb{N}}} \;=\;
\cDefDesX{g}{D \to \mathbb{N}}{
  \cEqX{g\;\mName{emptydir}}{0} \;\wedge\;
  \cForallBX{f}{F}{d}{D}{
    \cEqX{g\;(\mName{addfile}\;f\;d)}{g\;d + 1}}}$

\medskip
\noindent\textit{Remark.}
The function $\mName{size}\;d$ returns the number of file entries
in directory~$d$.

\medskip
\noindent\textbf{Thm\,3.} \quad $\cTotal{\mName{size}}$
\hfill ($\mName{size}$ is total).

\medskip
\noindent\textbf{Thm\,4.} \quad
$\cEqX{\mName{size}\;\mName{emptydir}}{0}$
\hfill (empty directory has size zero).

\medskip
\noindent\textbf{Thm\,5.} \quad
$\cForallBX{f}{F}{d}{D}{
  \cEqX{\mName{size}\;(\mName{addfile}\;f\;d)}{\mName{size}\;d + 1}}$
\hfill (addfile increments size).

\medskip
\noindent\textbf{Thm\,6.} \quad
$\cForallX{d}{D}{\mName{size}\;d \ge 0}$
\hfill (size is non-negative).

}
\end{dev-def}

\medskip
Definitions are validated and theorems are proved in
Appendix~\ref{app:proofs}.  Note that $\mName{contains}$ and
$\mName{size}$ are defined using definite descriptions
($\mathrm{I}$-terms): each definition asserts the existence of a
unique function satisfying the given recursion equations.  The
validations prove that these definite descriptions are defined
(i.e., the $\mathrm{I}$-terms denote values).

% ============================================================
\newpage
\section{Theory Morphism: Files to Directories}
\label{sec:morphism}
% ============================================================

We exhibit a theory morphism from $\FILETYPE$ to $\DIR$ to
demonstrate that individual files embed into the directory
structure.  Intuitively, a single file $f$ corresponds to the
one-element directory $\mName{addfile}\;f\;\mName{emptydir}$.

Since $\DIR$ is obtained from $\FILETYPE$ by a chain of
extensions ($\FILETYPE \leq \SECFILE \leq \DIR$), the constants
$\mkfile$, $\fname$, and $\fcontent$ and the base types $N$, $C$,
$F$ already belong to the language of $\DIR$.  We define a
\emph{non-trivial} morphism that reinterprets the file type $F$ as
the directory type $D$.

To define the accessor translations, we use the fact that
$\cInjB{\mName{addfile}}$ (Axiom~14) allows us to extract the
unique file from a single-entry directory.  Define:
\[
  \mName{extract}_{\cFunTyX{D}{F}} \;=\;
  \cFunAbsX{d}{D}{\cDefDesX{f}{F}{
    \cEqX{d}{\mName{addfile}\;f\;\mName{emptydir}}}}
\]
This definite description is defined on the image of the
embedding (single-entry directories) by Axiom~14.

\begin{theory-trans-def}
{Files to Directories}
{$\mName{FILE\text{-}to\text{-}DIR}$.}
{$\FILETYPE$.}
{$\DIR$.}
{%
\begin{itemize}
  \item[] $N \;\mapsto\; N$ \quad (names preserved)
  \item[] $C \;\mapsto\; C$ \quad (contents preserved)
  \item[] $F \;\mapsto\; D$
\end{itemize}
}
{%
\begin{itemize}
  \item[] $\mkfile_{N \to C \to F} \;\mapsto\;
    \cFunAbsX{n}{N}{\cFunAbsX{c}{C}{
      \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}}}$
  \item[] $\fname_{F \to N} \;\mapsto\;
    \cFunAbsX{d}{D}{\fname\;(\mName{extract}\;d)}$
  \item[] $\fcontent_{F \to C} \;\mapsto\;
    \cFunAbsX{d}{D}{\fcontent\;(\mName{extract}\;d)}$
\end{itemize}
}
\end{theory-trans-def}

\medskip
\noindent\textbf{Verification.}
We must show that the image of each $\FILETYPE$ axiom under this
morphism is valid in $\DIR$.  The key observation is that on
single-entry directories (those of the form
$\mName{addfile}\;f\;\mName{emptydir}$),
$\mName{extract}\;(\mName{addfile}\;f\;\mName{emptydir}) = f$
by the uniqueness of the definite description (using
$\cInjB{\mName{addfile}}$, Axiom~14).

\begin{enumerate}
  \item \textbf{TOTAL2($\mkfile$) maps to totality of the
    translation}: Since $\cTotalB{\mName{addfile}}$ (Axiom~13)
    and $\cTotalB{\mkfile}$ (Axiom~1),
    $\cFunAbsX{n}{N}{\cFunAbsX{c}{C}{
      \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}}}$
    is total.

  \item \textbf{INJ2($\mkfile$) maps to injectivity}: If
    $\mName{addfile}\;(\mkfile\;n_1\;c_1)\;\mName{emptydir}
    = \mName{addfile}\;(\mkfile\;n_2\;c_2)\;\mName{emptydir}$,
    then $\mkfile\;n_1\;c_1 = \mkfile\;n_2\;c_2$ by
    $\cInjB{\mName{addfile}}$ (Axiom~14), and $n_1 = n_2$,
    $c_1 = c_2$ by $\cInjB{\mkfile}$ (Axiom~2).

  \item \textbf{Accessor axioms (Axioms 3--4)}: For a
    single-entry directory
    $d = \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}$:
    $\fname\;(\mName{extract}\;d)
    = \fname\;(\mkfile\;n\;c) = n$ (Axiom~3),
    and similarly $\fcontent\;(\mName{extract}\;d) = c$ (Axiom~4).

  \item \textbf{Totality of accessors (Axioms 5--6)}: Follows
    from the totality of $\fname$, $\fcontent$, and the
    definedness of $\mName{extract}$ on single-entry directories.

  \item \textbf{Surjectivity (Axiom 7)}: For any single-entry
    directory $d$, $\mName{extract}\;d = f$ for some $f : F$,
    and the translated $\mkfile$ applied to $\fname\;f$ and
    $\fcontent\;f$ reconstructs $d$ by Axiom~7 of $\FILETYPE$.
\end{enumerate}

By the Morphism Theorem (Theorem~14.16 of~\cite{farmer2025}),
all theorems of $\FILETYPE$ transport to $\DIR$ under this
morphism.  This confirms that single files embed faithfully into
the directory structure.






% ============================================================
\newpage
\section{Theory Morphism: Files to Directories}
\label{sec:morphism}
% ============================================================

We exhibit a theory morphism from $\FILETYPE$ to $\DIR$ to
demonstrate that individual files embed into the directory
structure.  Intuitively, a single file $f$ corresponds to the
one-entry directory $\mName{addfile}\;f\;\mName{emptydir}$.

Since $\DIR$ extends $\SECFILE$, and $\SECFILE$ extends the
combined theory $\FILEPERM$, the base types $N$, $C$, $F$ and the
constants $\mkfile$, $\fname$, and $\fcontent$ from $\FILETYPE$
are all available in the language of $\DIR$.  To obtain a
well-formed theory translation in the sense of Chapter~14
of~\cite{farmer2025}, we do \emph{not} map the source base type
$F$ to the whole target type $D$.  Instead, we map $F$ to the
closed quasitype of \emph{single-entry directories}:
\[
  \zeta_F \;=\;
  \{\, d : D \mid \cForsomeX{f}{F}{
      \cEqX{d}{\mName{addfile}\;f\;\mName{emptydir}}}\,\}.
\]
This is a closed quasitype of $\DIR$, and it captures exactly the
image of the intended embedding of files into directories.

To define the accessor translations, we use the fact that
$\cInjB{\mName{addfile}}$ (Axiom~14) ensures that every
$d \in \zeta_F$ contains a unique file.  Define the extraction
function on $\zeta_F$ by:
\[
  \mName{extract}(d) \;=\;
  \cDefDesX{f}{F}{
    \cEqX{d}{\mName{addfile}\;f\;\mName{emptydir}}}.
\]
Because $d \in \zeta_F$, existence is guaranteed by definition of
$\zeta_F$, and uniqueness follows from $\cInjB{\mName{addfile}}$.

\begin{theory-trans-def}
{Files to Single-Entry Directories}
{$\mName{FILE\text{-}to\text{-}DIR}$.}
{$\FILETYPE$.}
{$\DIR$.}
{%
\begin{itemize}
  \item[] $N \;\mapsto\; N$ \quad (names preserved)
  \item[] $C \;\mapsto\; C$ \quad (contents preserved)
  \item[] $F \;\mapsto\; \zeta_F$
\end{itemize}
}
{%
\begin{itemize}
  \item[] $\mkfile_{N \to C \to F} \;\mapsto\;
    \cFunAbsX{n}{N}{\cFunAbsX{c}{C}{
      \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}}}$

  \item[] $\fname_{F \to N} \;\mapsto\;
    \cFunAbsX{d}{D}{\fname\;(\mName{extract}(d))}$

  \item[] $\fcontent_{F \to C} \;\mapsto\;
    \cFunAbsX{d}{D}{\fcontent\;(\mName{extract}(d))}$
\end{itemize}
}
\end{theory-trans-def}

\medskip
\noindent\textbf{Verification.}
We must show that the image of each axiom of $\FILETYPE$ under this
translation is valid in $\DIR$.

\begin{enumerate}
  \item \textbf{TOTAL2($\mkfile$).}
    The translated constructor
    $\cFunAbsX{n}{N}{\cFunAbsX{c}{C}{
        \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}}}$
    is total because $\cTotalB{\mkfile}$ (Axiom~1 of $\FILETYPE$)
    and $\cTotalB{\mName{addfile}}$ (Axiom~13 of $\DIR$) both hold.
    Its values lie in $\zeta_F$ by construction.

  \item \textbf{INJ2($\mkfile$).}
    If $\mName{addfile}\;(\mkfile\;n_1\;c_1)\;\mName{emptydir}
      =
      \mName{addfile}\;(\mkfile\;n_2\;c_2)\;\mName{emptydir}$,
    then $\mkfile\;n_1\;c_1 = \mkfile\;n_2\;c_2$ by
    $\cInjB{\mName{addfile}}$ (Axiom~14), and hence
    $n_1 = n_2$ and $c_1 = c_2$ by $\cInjB{\mkfile}$
    (Axiom~2 of $\FILETYPE$).

  \item \textbf{Accessor axioms (Axioms 3--4).}
    Let $d = \mName{addfile}\;(\mkfile\;n\;c)\;\mName{emptydir}$.
    Then $d \in \zeta_F$, so $\mName{extract}(d)$ is defined, and
    by uniqueness of the definite description, $\mName{extract}(d) = \mkfile\;n\;c$.
    Therefore $\fname\;(\mName{extract}(d))
      = \fname\;(\mkfile\;n\;c) = n$
    by Axiom~3 of $\FILETYPE$, and similarly 
    $\fcontent\;(\mName{extract}(d))
      = \fcontent\;(\mkfile\;n\;c) = c$
    by Axiom~4.

  \item \textbf{Totality of accessors (Axioms 5--6).}
    The translated accessors are total on $\zeta_F$ because for each
    $d \in \zeta_F$, the term $\mName{extract}(d)$ is defined, and
    $\fname$ and $\fcontent$ are total on $F$ by Axioms~5--6 of
    $\FILETYPE$.

  \item \textbf{Surjectivity (Axiom 7).}
    Let $d \in \zeta_F$.  Then, by definition of $\zeta_F$, there
    exists $f : F$ such that 
    $d = \mName{addfile}\;f\;\mName{emptydir}$.
    By definition of $\mName{extract}$ and uniqueness,
    $\mName{extract}(d)=f$.  Hence
      $\mName{addfile}\;(\mkfile\;(\fname\;f)\;(\fcontent\;f))\;\mName{emptydir}
      =
      \mName{addfile}\;f\;\mName{emptydir}
      = d$
    by Axiom~7 of $\FILETYPE$.
\end{enumerate}

Thus the translation is well formed.  By the Morphism Theorem
(Theorem~14.16 of~\cite{farmer2025}), all theorems of $\FILETYPE$
transport to $\DIR$ under $\mName{FILE\text{-}to\text{-}DIR}$.
This confirms that files embed faithfully into the substructure of
single-entry directories of $\DIR$.

% ============================================================
\newpage
\section{Examples}
\label{sec:examples}
% ============================================================

To illustrate the theories and development concretely, we consider
a small symbolic instance consisting of two files and a directory
containing them.  Let $n_1, n_2, n_3 : N$ and $c_1, c_2 : C$,
with $n_1 \neq n_2$, $n_3 \neq n_1$, and $n_3 \neq n_2$.

\medskip
\noindent\textbf{Example 8.1 (Secured files).}
Define
\[
  f_1 \;=\; \prot\;(\mkfile\;n_1\;c_1)\;\mName{rd},
  \qquad
  f_2 \;=\; \prot\;(\mkfile\;n_2\;c_2)\;\mName{wr}.
\]
By the axioms of $\FILETYPE$ and $\SECFILE$:
\[
  \fname\;f_1 = n_1,
  \qquad
  \fcontent\;f_1 = c_1,
  \qquad
  \perm\;f_1 = \mName{rd},
\]
and similarly
\[
  \fname\;f_2 = n_2,
  \qquad
  \fcontent\;f_2 = c_2,
  \qquad
  \perm\;f_2 = \mName{wr}.
\]
Thus $f_1$ and $f_2$ are files with the expected names, contents,
and permission levels.

\medskip
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
\textbf{File} & \textbf{Name / Content} & \textbf{Permission} \\
\hline
$f_1$ & $(n_1, c_1)$ & $\mName{rd}$ \\
$f_2$ & $(n_2, c_2)$ & $\mName{wr}$ \\
\hline
\end{tabular}
\end{center}

\medskip
\noindent\textbf{Example 8.2 (A directory with two entries).}
Define
\[
  d \;=\; \mName{addfile}\;f_1\;(\mName{addfile}\;f_2\;\mName{emptydir}).
\]
This is a flat directory with two entries.  By unfolding the
definition of $\mName{contains}$:
\[
  \mName{contains}\;n_1\;d = \cT,
  \qquad
  \mName{contains}\;n_2\;d = \cT,
  \qquad
  \mName{contains}\;n_3\;d = \cF.
\]
Indeed, $d$ contains files with names $n_1$ and $n_2$, but no file
with name $n_3$.  By the recursive definition of $\mName{size}$:
\[
  \mName{size}\;d
  = \mName{size}\;(\mName{addfile}\;f_1\;(\mName{addfile}\;f_2\;\mName{emptydir}))
  = \mName{size}\;(\mName{addfile}\;f_2\;\mName{emptydir}) + 1
  = (\mName{size}\;\mName{emptydir} + 1) + 1
  = 2.
\]
So the directory $d$ has exactly two file entries.

\medskip
\begin{center}
\begin{tabular}{|c|}
\hline
\textbf{Directory $d$} \\
\hline
$f_1$ \\
$f_2$ \\
\hline
\end{tabular}
\end{center}

\[
  \mName{contains}\;n_1\;d = \cT,
  \qquad
  \mName{contains}\;n_2\;d = \cT,
  \qquad
  \mName{contains}\;n_3\;d = \cF,
  \qquad
  \mName{size}\;d = 2.
\]

\medskip
\noindent\textbf{Example 8.3 (Embedding a file into a directory).}
The theory morphism $\mName{FILE\text{-}to\text{-}DIR}$ maps a file
to a one-entry directory.  In particular, the image of $f_1$ is
\[
  d_{f_1} \;=\; \mName{addfile}\;f_1\;\mName{emptydir}.
\]
Likewise,
\[
  d_{f_2} \;=\; \mName{addfile}\;f_2\;\mName{emptydir}.
\]
These are single-entry directories corresponding to the files
$f_1$ and $f_2$.  Intuitively, the morphism identifies an
individual file with the simplest directory that contains exactly
that file.

\medskip
\begin{center}
\begin{tabular}{ccc}
$f_1$ & $\longmapsto$ & $\mName{addfile}\;f_1\;\mName{emptydir}$ \\[4pt]
$f_2$ & $\longmapsto$ & $\mName{addfile}\;f_2\;\mName{emptydir}$
\end{tabular}
\end{center}

These examples illustrate the intended behaviour of the six modules:
$\PERM$ supplies the permission levels, $\FILETYPE$ constructs files,
$\SECFILE$ assigns permissions to them, $\DIR$ builds flat
directories from files, $\DIRDEV$ defines directory operations, and
$\mName{FILE\text{-}to\text{-}DIR}$ embeds files into the directory
structure.

% ============================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================

We have formalized a simple file system as a development graph in
Alonzo consisting of six modules:

\begin{enumerate}
  \item $\PERM$: an enumeration inductive type specifying four
    permission levels.
  \item $\FILETYPE$: a theory definition specifying files as
    name--content pairs with a constructor and accessors.
  \item $\SECFILE$: a theory extension combining files with
    permissions via $\prot$ and $\perm$.
  \item $\DIR$: a theory extension introducing directories as a
    recursive inductive type over secured files.
  \item $\DIRDEV$: a development definition with two defined
    operations ($\mName{contains}$, $\mName{size}$) and six
    theorems, including totality proofs and algebraic properties.
  \item $\mName{FILE\text{-}to\text{-}DIR}$: a theory morphism
    showing that files embed into directories, validated by the
    Morphism Theorem (Theorem~14.16 of~\cite{farmer2025}).
\end{enumerate}

The development forms a genuine graph (not a linear chain):
$\PERM$ and $\FILETYPE$ are independent base theories that merge
into $\SECFILE$, which is then extended to $\DIR$.


% ============================================================
\appendix
\newpage
\section{Proofs and Validations for DIR-DEV}
\label{app:proofs}
% ============================================================

\subsection{Validation of Def\,1 ($\mName{contains}$)}

We must show that the definite description
$\cDefDesX{g}{N \to D \to \cB}{\ldots}$
denotes a value, i.e., there exists a unique function
$g : N \to D \to \cB$ satisfying:
\begin{align}
  \cForallX{n}{N}{\cEqX{g\;n\;\mName{emptydir}}{\cF}} \tag{R1} \\
  \cForallCX{n}{N}{f}{F}{d}{D}{
    \cEqX{g\;n\;(\mName{addfile}\;f\;d)}{
      (\cEqX{\fname\;f}{n}) \vee g\;n\;d}}
    \tag{R2}
\end{align}

\textbf{Existence.}
Define $g$ by: for fixed $n : N$, define $g\;n$ by structural
recursion on the second argument $d : D$.
By the structural induction principle (Axiom~16 of $\DIR$),
every $d : D$ is either $\mName{emptydir}$ or
$\mName{addfile}\;f\;d'$ for unique $f : F$ and $d' : D$.  Define:
\[
  g\;n\;d \;=\;
  \begin{cases}
    \cF & \text{if } d = \mName{emptydir}, \\
    (\cEqX{\fname\;f}{n}) \vee g\;n\;d'
      & \text{if } d = \mName{addfile}\;f\;d'.
  \end{cases}
\]
This is well-defined by structural recursion (each recursive call
is on a structurally smaller second argument).  The expression
$(\cEqX{\fname\;f}{n}) \vee g\;n\;d'$ is defined since $\fname$ is
total (Axiom~5 of $\FILETYPE$), equality is always defined, and
$\vee$ is total.  Hence $g$ satisfies (R1) and (R2).

\textbf{Uniqueness.}
If $g$ and $h$ both satisfy (R1) and (R2), then for any fixed
$n : N$, by structural induction on $d$:
\begin{itemize}
  \item \textbf{Base:}
    $g\;n\;\mName{emptydir} = \cF = h\;n\;\mName{emptydir}$.
  \item \textbf{Step:} If $g\;n\;d = h\;n\;d$, then
    $g\;n\;(\mName{addfile}\;f\;d)
    = (\cEqX{\fname\;f}{n}) \vee g\;n\;d
    = (\cEqX{\fname\;f}{n}) \vee h\;n\;d
    = h\;n\;(\mName{addfile}\;f\;d)$.
\end{itemize}
Hence $g = h$. \qed

\subsection{Thm\,1: $\cTotalB{\mName{contains}}$}

\textbf{Proof.}
We show $\mName{contains}\;n\;d\!\downarrow$ for all $n : N$,
$d : D$ by structural induction on $d$.

\textbf{Base case:}
$\mName{contains}\;n\;\mName{emptydir} = \cF\!\downarrow$.

\textbf{Inductive step:} Assume
$\mName{contains}\;n\;d\!\downarrow$.  Then
$\mName{contains}\;n\;(\mName{addfile}\;f\;d)
= (\cEqX{\fname\;f}{n}) \vee \mName{contains}\;n\;d$.
Since $\fname\;f\!\downarrow$ (Axiom~5), equality is defined,
$\mName{contains}\;n\;d\!\downarrow$ (IH), and $\vee$ is total,
the result is defined.

By structural induction (Axiom~16), $\cTotalB{\mName{contains}}$.
\qed

\subsection{Thm\,2:
  $\cForallX{n}{N}{\cEqX{\mName{contains}\;n\;\mName{emptydir}}{\cF}}$}

Immediate from the definition of $\mName{contains}$ (equation R1).
\qed

\subsection{Validation of Def\,2 ($\mName{size}$)}

We must show that $\cDefDesX{g}{D \to \mathbb{N}}{\ldots}$ denotes
a value.

\textbf{Existence.}
Define $g$ by structural recursion on $d$:
$g\;\mName{emptydir} = 0$ and
$g\;(\mName{addfile}\;f\;d) = g\;d + 1$.
This is well-defined since the recursive call is on the
structurally smaller argument $d$, and $+ 1$ is total on
$\mathbb{N}$.

\textbf{Uniqueness.}
By structural induction: if $g$ and $h$ both satisfy the
recursion, then $g\;\mName{emptydir} = 0 = h\;\mName{emptydir}$,
and if $g\;d = h\;d$ then
$g\;(\mName{addfile}\;f\;d) = g\;d + 1 = h\;d + 1
= h\;(\mName{addfile}\;f\;d)$.
Hence $g = h$. \qed

\subsection{Thm\,3: $\cTotal{\mName{size}}$}

\textbf{Proof.}
By structural induction on $d$.

\textbf{Base:} $\mName{size}\;\mName{emptydir} = 0\!\downarrow$.

\textbf{Step:} Assume $\mName{size}\;d\!\downarrow$.  Then
$\mName{size}\;(\mName{addfile}\;f\;d) = \mName{size}\;d + 1$.
Since $\mName{size}\;d\!\downarrow$ (IH) and $+ 1$ is total on
$\mathbb{N}$, the result is defined. \qed

\subsection{Thm\,4:
  $\cEqX{\mName{size}\;\mName{emptydir}}{0}$}

Immediate from the definition of $\mName{size}$. \qed

\subsection{Thm\,5:
  $\cForallBX{f}{F}{d}{D}{
    \cEqX{\mName{size}\;(\mName{addfile}\;f\;d)}{\mName{size}\;d + 1}}$}

Immediate from the definition of $\mName{size}$. \qed

\subsection{Thm\,6:
  $\cForallX{d}{D}{\mName{size}\;d \ge 0}$}

\textbf{Proof.}
By structural induction on $d$.

\textbf{Base:} $\mName{size}\;\mName{emptydir} = 0 \ge 0$.

\textbf{Step:} Assume $\mName{size}\;d \ge 0$.  Then
$\mName{size}\;(\mName{addfile}\;f\;d) = \mName{size}\;d + 1
\ge 0 + 1 = 1 \ge 0$. \qed

\begin{thebibliography}{9}

\bibitem{farmer2025}
  W.~M. Farmer,
  \emph{Simple Type Theory: A Practical Logic for Expressing and
  Reasoning About Mathematical Ideas},
  2nd ed., Birkh{\"a}user, 2025.

\bibitem{monoids2025}
  W.~M. Farmer and D.~Y. Zvigelsky,
  ``Monoid Theory in Alonzo: A Little Theories Formalization in
  Simple Type Theory,''
  \emph{Journal of Applied Logics}, vol.~12, no.~7, 2025.

\end{thebibliography}

\end{document}
